---
title: "Stats Project"
author: "Miguel/Antonia"
date: "2022-12-19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages and set working directory
```{r}
library(glmnet)
library(gridExtra)
#source(file.path("./SIM_project/helper_functions/routines_seminar1.R"))
library("readxl")
library(lme4)
library(arm)
library(tidyverse)
```

Set seed
```{r}
set.seed(1234)
```

Load data
```{r load, include=FALSE}
f22 <- read_csv("./data/players_22.csv")
f21 <- read_csv("./data/players_21.csv")
f20 <- read_csv("./data/players_20.csv")
f19 <- read_csv("./data/players_19.csv")
f18 <- read_csv("./data/players_18.csv")
f17 <- read_csv("./data/players_17.csv")
f16 <- read_csv("./data/players_16.csv")
f15 <- read_csv("./data/players_15.csv")
```

Remove some unnecessary columns
```{r}
f22 <- f22 %>% dplyr::select(-c("player_face_url", "club_logo_url", "club_flag_url", 
                         "nation_logo_url", "nation_flag_url", "player_url",
                         "real_face"))

f21 <- f21 %>% dplyr::select(-c("player_face_url", "club_logo_url", "club_flag_url", 
                         "nation_logo_url", "nation_flag_url", "player_url",
                         "real_face"))

f20 <- f20 %>% dplyr::select(-c("player_face_url", "club_logo_url", "club_flag_url", 
                         "nation_logo_url", "nation_flag_url", "player_url",
                         "real_face"))

f19 <- f19 %>% dplyr::select(-c("player_face_url", "club_logo_url", "club_flag_url", 
                         "nation_logo_url", "nation_flag_url", "player_url",
                         "real_face"))

f18 <- f18 %>% dplyr::select(-c("player_face_url", "club_logo_url", "club_flag_url", 
                         "nation_logo_url", "nation_flag_url", "player_url",
                         "real_face"))

f17 <- f17 %>% dplyr::select(-c("player_face_url", "club_logo_url", "club_flag_url", 
                         "nation_logo_url", "nation_flag_url", "player_url",
                         "real_face"))

f16 <- f16 %>% dplyr::select(-c("player_face_url", "club_logo_url", "club_flag_url", 
                         "nation_logo_url", "nation_flag_url", "player_url",
                         "real_face"))

f15 <- f15 %>% dplyr::select(-c("player_face_url", "club_logo_url", "club_flag_url", 
                         "nation_logo_url", "nation_flag_url", "player_url",
                         "real_face"))
```


### TIMELINE ###
# Season Starts     (Fall 2020)
# Season Ends         (Spring 2021)
# Game (FIFA 2022) is made and players are rated (Summer 2021) 
# New Teams (Summer 2021) (These come from the 2022 edition of the game because they are from the 2021-2022 season.)
# Predictions are Made for next season (Summer 2021)
# New Season (2021 -> 2022)

# So, we are going to use the FIFA 2022 ratings to predict the 2021-2022 season with the FIFA 2022 rosters. 
# The 2021-2022 points data goes with FIFA 2022, since that is what we are trying to predict.

Import REAL end of season info (total number of points, wins, losses, etc. for each team)
```{r}
table22 <- read_excel("./data/LaLiga_results.xlsx", sheet = "2021-2022")
table21 <- read_excel("./data/LaLiga_results.xlsx", sheet = "2020-2021")
table20 <- read_excel("./data/LaLiga_results.xlsx", sheet = "2019-2020")
table19 <- read_excel("./data/LaLiga_results.xlsx", sheet = "2018-2019")
table18 <- read_excel("./data/LaLiga_results.xlsx", sheet = "2017-2018")
table17 <- read_excel("./data/LaLiga_results.xlsx", sheet = "2016-2017")
table16 <- read_excel("./data/LaLiga_results.xlsx", sheet = "2015-2016")
table15 <- read_excel("./data/LaLiga_results.xlsx", sheet = "2014-2015")
```

View data
```{r}
f22
colnames(f22)
```

Filter for only Spain Primera Division
```{r}
f22_sp <- f22 %>% filter(league_name=='Spain Primera Division')
f21_sp <- f21 %>% filter(league_name=='Spain Primera Division')
f20_sp <- f20 %>% filter(league_name=='Spain Primera Division')
f19_sp <- f19 %>% filter(league_name=='Spain Primera Division')
f18_sp <- f18 %>% filter(league_name=='Spain Primera Division')
f17_sp <- f17 %>% filter(league_name=='Spain Primera Division')
f16_sp <- f16 %>% filter(league_name=='Spain Primera Division')
f15_sp <- f15 %>% filter(league_name=='Spain Primera Division')
```

Pick only the columns that we want.
```{r}
cols_to_choose <- c("overall", "potential", "value_eur", "wage_eur", "age", #"dob", 
                    "height_cm", "weight_kg",  "club_name", #"club_team_id",
                    #"league_level", 
                    "preferred_foot", "weak_foot", "skill_moves", "international_reputation", #"work_rate", #"body_type"                  
                    "release_clause_eur", #"player_tags", #"player_traits", 
                    "pace", 
                    "club_position",
                    "shooting", "passing", "dribbling", "defending",                  
                    "physic", "attacking_crossing", "attacking_finishing", "attacking_heading_accuracy", 
                    "attacking_short_passing", "attacking_volleys", "skill_dribbling", "skill_curve",                
                    "skill_fk_accuracy", "skill_long_passing", "skill_ball_control", "movement_acceleration",      
                    "movement_sprint_speed", "movement_agility", "movement_reactions", "movement_balance",           
                    "power_shot_power", "power_jumping", "power_stamina", "power_strength",             
                    "power_long_shots", "mentality_aggression", "mentality_interceptions", "mentality_positioning",      
                    "mentality_vision", "mentality_penalties", "mentality_composure", "defending_marking_awareness",
                    "defending_standing_tackle", "defending_sliding_tackle", "goalkeeping_diving", "goalkeeping_handling",
                    "goalkeeping_kicking", "goalkeeping_positioning", "goalkeeping_reflexes", "goalkeeping_speed" 
                    #"ls", "st", "rs","lw", 
                    #"lf", "cf", "rf", "rw",
                    #"lam", "cam", "ram", "lm", 
                    #"lcm", "cm", "rcm", "rm",                       
                    #"lwb", "ldm", "cdm", "rdm",                        
                    #"rwb", "lb", "lcb", "cb",                        
                    #"rcb", "rb", "gk" 
                    ) 
```

Make the column selection.
```{r}
f22_sp <- f22_sp[cols_to_choose]
f21_sp <- f21_sp[cols_to_choose]
f20_sp <- f20_sp[cols_to_choose]
f19_sp <- f19_sp[cols_to_choose]
f18_sp <- f18_sp[cols_to_choose]
f17_sp <- f17_sp[cols_to_choose]
f16_sp <- f16_sp[cols_to_choose]
f15_sp <- f15_sp[cols_to_choose]
```

Make simplified position: GK/Defense/Midfield/Forward/Reserve
```{r}
get_pos <- function(setx) {
  setx$pos <- c(rep("POS", nrow(setx)))
  defc <- setx$club_position %in% c("LCB", "LB", "RB", "RCB", "CB", "RWB", "LWB")
  midc <- setx$club_position %in% c("CDM","LCM","RCM","RM","CAM","LW","LM","RW","RDM","LDM","CM")
  forc <- setx$club_position %in% c("RS", "ST", "LS", "CF", "LF", "RF")
  subc <- setx$club_position %in% c("SUB", "RES")
  gkc <- setx$club_position %in% c("GK")
  setx$pos[defc] <- "defender"
  setx$pos[midc] <- "midfielder"
  setx$pos[forc] <- "forward"
  setx$pos[subc] <- "sub/res"
  setx$pos[gkc] <- "goalkeeper"
  return(setx)
}

f22_sp <- get_pos(f22_sp)
f21_sp <- get_pos(f21_sp)
f20_sp <- get_pos(f20_sp)
f19_sp <- get_pos(f19_sp)
f18_sp <- get_pos(f18_sp)
f17_sp <- get_pos(f17_sp)
f16_sp <- get_pos(f16_sp)
f15_sp <- get_pos(f15_sp)
```

Preprocessing of data
```{r}
# Replace NAs by 0
f22_sp[is.na(f22_sp)] <- 0
f21_sp[is.na(f21_sp)] <- 0
f20_sp[is.na(f20_sp)] <- 0
f19_sp[is.na(f19_sp)] <- 0
f18_sp[is.na(f18_sp)] <- 0
f17_sp[is.na(f17_sp)] <- 0
f16_sp[is.na(f16_sp)] <- 0
f15_sp[is.na(f15_sp)] <- 0
# Group dataframes by team.
f22_gp <- f22_sp %>% group_by(club_name) %>% summarise(across(where(is.numeric), mean))
f21_gp <- f21_sp %>% group_by(club_name) %>% summarise(across(where(is.numeric), mean))
f20_gp <- f20_sp %>% group_by(club_name) %>% summarise(across(where(is.numeric), mean))
f19_gp <- f19_sp %>% group_by(club_name) %>% summarise(across(where(is.numeric), mean))
f18_gp <- f18_sp %>% group_by(club_name) %>% summarise(across(where(is.numeric), mean))
f17_gp <- f17_sp %>% group_by(club_name) %>% summarise(across(where(is.numeric), mean))
f16_gp <- f16_sp %>% group_by(club_name) %>% summarise(across(where(is.numeric), mean))
f15_gp <- f15_sp %>% group_by(club_name) %>% summarise(across(where(is.numeric), mean))
# Join dataframes
f22_gp <- left_join(f22_gp, table22, by = c("club_name"="Equipo"))
f21_gp <- left_join(f21_gp, table21, by = c("club_name"="Equipo"))
f20_gp <- left_join(f20_gp, table20, by = c("club_name"="Equipo"))
f19_gp <- left_join(f19_gp, table19, by = c("club_name"="Equipo"))
f18_gp <- left_join(f18_gp, table18, by = c("club_name"="Equipo"))
f17_gp <- left_join(f17_gp, table17, by = c("club_name"="Equipo"))
f16_gp <- left_join(f16_gp, table16, by = c("club_name"="Equipo"))
f15_gp <- left_join(f15_gp, table15, by = c("club_name"="Equipo"))
# # Check for NAs in dataframe.
# for(i in 1:length(cols_to_choose)){
#   print(sum(is.na(f22_gp[i])))
# }
```

Add a multiplier on points for second division teams going to first division.
```{r}
# second_div_mult <- 0.5
```

Create new dataframe
```{r}
new_df <- bind_rows(f21_gp, f20_gp, f19_gp, f18_gp, f17_gp, f16_gp, f15_gp)
```

Remove some columns
```{r}
colnames(new_df)
cols_to_drop <- c( "Pos.", "PJ", "G", "E", "P", "GF", "GC", "Dif.", "Des", "Asc")
new_df <- new_df[, !colnames(new_df) %in% cols_to_drop]
f22_gp <- f22_gp[, !colnames(f22_gp) %in% cols_to_drop]
```

TRAIN
```{r}
y <- new_df$Pts.
x <- dplyr::select(new_df, -c("Pts.", "club_name") ) %>% as.matrix() %>% scale()
```

TEST 
#### NEEDS TO BE SAME SCALE!!
```{r}
y_te <- f22_gp[c("club_name","Pts.")]
x_te <- dplyr::select(f22_gp, -c("Pts.", "club_name") ) %>% as.matrix() %>% scale()
```
# -----------------------------------
# ----------- LASSO MODEL -----------
# -----------------------------------
```{r}
fit.lasso= cv.glmnet(x=x, y=y, family = "gaussian", nfolds = 5, alpha=1, type.measure = "mse")

# model
lasso.err <- assess.glmnet(fit.lasso, newx=x, newy=y, family='gaussian')
lasso.err 

plot(fit.lasso$glmnet.fit, xvar='lambda')
plot(fit.lasso)

# try model w min value of lambda
b.lasso= as.vector(coef(fit.lasso, s='lambda.min'))
betas = data.frame("word" = c("intercept", colnames(x)), "beta_hat"= round(b.lasso, 5))
betas %>% arrange(desc(abs(beta_hat)))
```

```{r diff lambda}
# not as good but just to try something else
b.lasso= as.vector(coef(fit.lasso, s=0.1))
betas = data.frame("word" = c("intercept", colnames(x)), "beta_hat"= round(b.lasso, 5))
betas %>% arrange(desc(abs(beta_hat)))

# Test base model on 22 dataset
predictions <- predict(fit.lasso, newx = x_te)
y_te$pred <-predictions
y_te <- y_te %>% arrange(desc(Pts.))
```


```{r}
# Test other model on 22 dataset
predictions <- predict(fit.lasso, newx = x_te)
y_te$pred <-predictions
y_te <- y_te %>% arrange(desc(Pts.))
# For dataset
# compute averages by player
# calculate month of birth for player
# modify goalkeeper stats to 0 for 

# For model
# different values of lambda
```
# -----------------------------------
# ------ HIERARCHICAL MODEL ---------
# -----------------------------------
```{r}
fp22_hm <- left_join(f22_sp, table22, by = c("club_name"="Equipo"))
fp21_hm <- left_join(f21_sp, table21, by = c("club_name"="Equipo"))
fp20_hm <- left_join(f20_sp, table20, by = c("club_name"="Equipo"))
fp19_hm <- left_join(f19_sp, table19, by = c("club_name"="Equipo"))
fp18_hm <- left_join(f18_sp, table18, by = c("club_name"="Equipo"))
fp17_hm <- left_join(f17_sp, table17, by = c("club_name"="Equipo"))
fp16_hm <- left_join(f16_sp, table16, by = c("club_name"="Equipo"))
fp15_hm <- left_join(f15_sp, table15, by = c("club_name"="Equipo"))


cols_to_drop <- c( "Pos.", "PJ", "G", "E", "P", "GF", "GC", "Dif.", "Des", "Asc", "league_level")

y_te2 <- fp22_hm[c("Pts.","club_name")]
hm_data <-bind_rows(fp21_hm, fp20_hm, fp19_hm, fp18_hm, fp17_hm, fp16_hm, fp15_hm)
hm_data<- hm_data[, !colnames(hm_data) %in% cols_to_drop] 
fp22_hm<- fp22_hm[, !colnames(fp22_hm) %in% cols_to_drop] 

# fp22_hm[] <- lapply(fp22_hm, function(x) if(is.numeric(x)){
#   scale(x, center=TRUE, scale=TRUE)
#   } else x)

# Check for NAs in dataframe.
for(i in 1:length(colnames(fp22_hm))){
  print(sum(is.na(fp22_hm[i])))
}
# fp22_hm$ran <- sample(-10:10, 633, rep = TRUE)
# fp22_hm$ranpts <- fp22_hm$Pts.+ fp22_hm$ran

# later add right/left foot as binary variable

hm_datasc <- hm_data %>% 
  dplyr::select(-c(preferred_foot, club_name, Pts., club_position, pos)) %>% 
  scale() %>% 
  data.frame () %>% 
  mutate_if(is.character, as.numeric)
hm_datasc <- cbind(hm_datasc, 'club_name'= hm_data$club_name, 
                   'Pts.'=hm_data$Pts., 
                   'club_position'=hm_data$club_position, 
                   'pos'=hm_data$pos) 

fp22_hm_sc <- fp22_hm %>% 
  dplyr::select(-c(preferred_foot, club_name, Pts., club_position, pos)) %>% 
  scale() %>% 
  data.frame () %>% 
  mutate_if(is.character, as.numeric)
fp22_hm_sc <- cbind(fp22_hm_sc, 
                    'club_name'= fp22_hm$club_name, 
                    'Pts.'=fp22_hm$Pts., 
                    'club_position'=fp22_hm$club_position,
                    'pos'=fp22_hm$pos) 


hm_datasc <- data.frame(hm_datasc)
fp22_hm_sc <- data.frame(fp22_hm_sc)

str(hm_datasc)


ML_ex <- lmer(Pts. ~age + overall + international_reputation  + (1|pos), data=hm_datasc,  REML = FALSE) 
pts_pred<-predict(ML_ex, newdata=fp22_hm_sc)
ML_ex
d <- f22 %>% filter(league_name=='Spain Primera Division')
compare <- cbind(y_te2, pts_pred, d$short_name)
compare <- compare %>% mutate("diff" = Pts.-pts_pred)
sd = mean((compare$diff)^2)^0.5
sd
#[1] 12.16879 age + overall + international_reputation  + (1|pos)
#[1] 7.522987  age + overall + international_reputation  + (1|club_name)
#[1] 7.4702 age + overall + international_reputation  + (1|club_name:pos)

compare
```

Try to predict player value from other data.
```{r}
hm_datasc["value_eur"] <- hm_data$value_eur
fp22_hm_sc["value_eur"] <- fp22_hm$value_eur

ML_ex_o <- lmer(value_eur ~  age + overall + international_reputation + wage_eur + (1|club_position), data=hm_datasc,  REML = FALSE)
v_pred <-predict(ML_ex_o, newdata=fp22_hm_sc)
ML_ex_o

compare_o <- cbind(y_te2, v_pred, d[c("short_name", "overall", "value_eur")])
compare_o <- compare_o %>% mutate("diff" = round(value_eur-v_pred))
mean((compare_o$diff)^2)^0.5
# 9026972
```

Try to predict player wage from data
```{r}
hm_datasc["wage_eur"] <- hm_data$wage_eur
fp22_hm_sc["wage_eur"] <- fp22_hm$wage_eur

ML_ex_o <- lmer(wage_eur ~  age + overall + international_reputation + (1|club_name:pos), data=hm_datasc,  REML = FALSE) 
w_pred <-predict(ML_ex_o, newdata=fp22_hm_sc)
ML_ex_o

compare_w <- cbind(y_te2, v_pred, d[c("short_name", "overall", "wage_eur")])
compare_w <- compare_w %>% mutate("diff" = round(wage_eur-w_pred))
mean((compare_w$diff)^2)^0.5
# 24548.54 (1|pos) + (1|club_name)
# 24591.14 (1|club_name)
# 23680.58 (1|club_position) + (1|club_name)
# 21600.1(1|club_name:pos)
# 21468.42 (1|club_name) + (1|club_name:pos)


mean(hm_data$wage_eur)


```

# -----------------------------------
# ---- BAYESIAN REGRESSION MODEL ----
# -----------------------------------

```{r}
# Using scaled_new_df dataframe for Bayesian regression model
scaled_new_df <- bind_cols(x, y)
names(scaled_new_df)[names(scaled_new_df) == '...53'] <- 'Pts.'
```

Fitting the model
```{r}
library(rstanarm)

# Fit the model:
bayes_reg_model <- rstanarm::stan_glm(Pts. ~ ., family = gaussian(), data=scaled_new_df)
```

View a summary of the model:
```{r}
summary(bayes_reg_model)
```

To obtain summary of medians
```{r}
bayes_reg_model
```

To obtain prediction intervals:
```{r}
m= coef(bayes_reg_model)
q= posterior_interval(bayes_reg_model)[1:length(m),]
b.intervals= round(cbind(m, q), 3)
b.intervals
```

Making predictions:
```{r}
# Making predictions:
x_te_df <- as.data.frame(x_te)
predictions_bayes <- predict(bayes_reg_model, newdata = x_te_df)

pred_bayes_compare <- cbind(y_te, predictions_bayes)
pred_bayes_compare
```

Mean squared error to assess accuracy of Bayesian Regression
```{r}
mse_bayes = mean((pred_bayes_compare$Pts. - pred_bayes_compare$predictions_bayes)^2)
mse_bayes
```

Root mean squared error to assess accuracy of Bayesian Regression
```{r}
root_mse_bayes = sqrt(mean((pred_bayes_compare$Pts. - pred_bayes_compare$predictions_bayes)^2))
root_mse_bayes
```

Correlation between predicted and actual values of testing set
```{r}
cor(pred_bayes_compare$Pts., pred_bayes_compare$predictions_bayes)
```


# -----------------------------------
# ---- BAYESIAN MODEL SELECTION -----
# -----------------------------------

Fitting the model
```{r}
library(mombf)
# Fit Bayesian model selection model
fit.bayesmodelselection <- modelSelection(Pts. ~ ., data=scaled_new_df, priorCoef=zellnerprior(taustd=1), priorDelta=modelbbprior(1,1))
```

Viewing summary of bayesian model selection
```{r}
head(postProb(fit.bayesmodelselection),10)
```


# -----------------------------------
# --------- VISUALIZATIONS ----------
# -----------------------------------

```{r}
colnames(scaled_new_df)
```

```{r}
scaled_new_df[1, ]
```

```{r pressure, echo=FALSE}
library(ggplot2)
ggplot(data=new_df, aes(Pts.)) + geom_histogram(bins = 20) + ggtitle("Distribution of Points Scored")
#hist(new_df$Pts.)
```

```{r pressure1, echo=FALSE}
ggplot(data=new_df, aes(Pts.)) + geom_density()
```