---
title: "Stats Project"
output: html_document
date: "2022-12-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages and set working directory
```{r}
library(tidyverse)
library(glmnet)
library(gridExtra)
setwd("~/Desktop")
library("readxl")
library(lme4)
library(arm)
#PATH= "./stats_project/SIM_project-main"
```

Set seed
```{r}
set.seed(1234)
```

Load data
```{r}
f22 <- read_csv("./stats_project/SIM_project-main/data/players_22.csv")
f21 <- read_csv("./stats_project/SIM_project-main//data/players_21.csv")
f20 <- read_csv("./stats_project/SIM_project-main//data/players_20.csv")
f19 <- read_csv("./stats_project/SIM_project-main//data/players_19.csv")
f18 <- read_csv("./stats_project/SIM_project-main//data/players_18.csv")
f17 <- read_csv("./stats_project/SIM_project-main//data/players_17.csv")
f16 <- read_csv("./stats_project/SIM_project-main//data/players_16.csv")
f15 <- read_csv("./stats_project/SIM_project-main//data/players_15.csv")
```

Remove some unnecessary columns
```{r}
f22 <- f22 %>% select(-c("player_face_url", "club_logo_url", "club_flag_url", 
                         "nation_logo_url", "nation_flag_url", "player_url",
                         "real_face"))

f21 <- f21 %>% select(-c("player_face_url", "club_logo_url", "club_flag_url", 
                         "nation_logo_url", "nation_flag_url", "player_url",
                         "real_face"))

f20 <- f20 %>% select(-c("player_face_url", "club_logo_url", "club_flag_url", 
                         "nation_logo_url", "nation_flag_url", "player_url",
                         "real_face"))

f19 <- f19 %>% select(-c("player_face_url", "club_logo_url", "club_flag_url", 
                         "nation_logo_url", "nation_flag_url", "player_url",
                         "real_face"))

f18 <- f18 %>% select(-c("player_face_url", "club_logo_url", "club_flag_url", 
                         "nation_logo_url", "nation_flag_url", "player_url",
                         "real_face"))

f17 <- f17 %>% select(-c("player_face_url", "club_logo_url", "club_flag_url", 
                         "nation_logo_url", "nation_flag_url", "player_url",
                         "real_face"))

f16 <- f16 %>% select(-c("player_face_url", "club_logo_url", "club_flag_url", 
                         "nation_logo_url", "nation_flag_url", "player_url",
                         "real_face"))

f15 <- f15 %>% select(-c("player_face_url", "club_logo_url", "club_flag_url", 
                         "nation_logo_url", "nation_flag_url", "player_url",
                         "real_face"))
```

Check they all have the same length
```{r}
length(colnames(f22))
length(colnames(f21))
length(colnames(f20))
length(colnames(f19))
length(colnames(f18))
length(colnames(f17))
length(colnames(f16))
length(colnames(f15))
```

### TIMELINE ###
# Season Starts     (Fall 2020)
# Season Ends         (Spring 2021)
# Game (FIFA 2022) is made and players are rated (Summer 2021) 
# New Teams (Summer 2021) (These come from the 2022 edition of the game because they are from the 2021-2022 season.)
# Predictions are Made for next season (Summer 2021)
# New Season (2021 -> 2022)

# So, we are going to use the FIFA 2022 ratings to predict the 2021-2022 season with the FIFA 2022 rosters. 
# The 2021-2022 points data goes with FIFA 2022, since that is what we are trying to predict.

Import REAL end of season info (total number of points, wins, losses, etc. for each team)
```{r}
table22 <- read_excel("./stats_project/SIM_project-main/data/LaLiga_results.xlsx", sheet = "2021-2022")
table21 <- read_excel("./stats_project/SIM_project-main/data/LaLiga_results.xlsx", sheet = "2020-2021")
table20 <- read_excel("./stats_project/SIM_project-main/data/LaLiga_results.xlsx", sheet = "2019-2020")
table19 <- read_excel("./stats_project/SIM_project-main/data/LaLiga_results.xlsx", sheet = "2018-2019")
table18 <- read_excel("./stats_project/SIM_project-main/data/LaLiga_results.xlsx", sheet = "2017-2018")
table17 <- read_excel("./stats_project/SIM_project-main/data/LaLiga_results.xlsx", sheet = "2016-2017")
table16 <- read_excel("./stats_project/SIM_project-main/data/LaLiga_results.xlsx", sheet = "2015-2016")
table15 <- read_excel("./stats_project/SIM_project-main/data/LaLiga_results.xlsx", sheet = "2014-2015")
```

View data
```{r}
f22
colnames(f22)
```

Filter for only Spain Primera Division
```{r}
f22_sp <- f22 %>% filter(league_name=='Spain Primera Division')
f21_sp <- f21 %>% filter(league_name=='Spain Primera Division')
f20_sp <- f20 %>% filter(league_name=='Spain Primera Division')
f19_sp <- f19 %>% filter(league_name=='Spain Primera Division')
f18_sp <- f18 %>% filter(league_name=='Spain Primera Division')
f17_sp <- f17 %>% filter(league_name=='Spain Primera Division')
f16_sp <- f16 %>% filter(league_name=='Spain Primera Division')
f15_sp <- f15 %>% filter(league_name=='Spain Primera Division')
```

Pick only the columns that we want.
```{r}
cols_to_choose <- c("overall", "potential", "value_eur", "wage_eur", "age", #"dob", 
                    "height_cm", "weight_kg",  "club_name", #"club_team_id",
                    "league_level", "preferred_foot", "weak_foot", "skill_moves", "international_reputation", #"work_rate", #"body_type"                  
                    "release_clause_eur", #"player_tags", #"player_traits", 
                    "pace", 
                    "shooting", "passing", "dribbling", "defending",                  
                    "physic", "attacking_crossing", "attacking_finishing", "attacking_heading_accuracy", 
                    "attacking_short_passing", "attacking_volleys", "skill_dribbling", "skill_curve",                
                    "skill_fk_accuracy", "skill_long_passing", "skill_ball_control", "movement_acceleration",      
                    "movement_sprint_speed", "movement_agility", "movement_reactions", "movement_balance",           
                    "power_shot_power", "power_jumping", "power_stamina", "power_strength",             
                    "power_long_shots", "mentality_aggression", "mentality_interceptions", "mentality_positioning",      
                    "mentality_vision", "mentality_penalties", "mentality_composure", "defending_marking_awareness",
                    "defending_standing_tackle", "defending_sliding_tackle", "goalkeeping_diving", "goalkeeping_handling",
                    "goalkeeping_kicking", "goalkeeping_positioning", "goalkeeping_reflexes", "goalkeeping_speed" 
                    #"ls", "st", "rs","lw", 
                    #"lf", "cf", "rf", "rw",
                    #"lam", "cam", "ram", "lm", 
                    #"lcm", "cm", "rcm", "rm",                       
                    #"lwb", "ldm", "cdm", "rdm",                        
                    #"rwb", "lb", "lcb", "cb",                        
                    #"rcb", "rb", "gk" 
                    ) 
```

Make the column selection.
```{r}
f22_sp <- f22_sp[cols_to_choose]
f21_sp <- f21_sp[cols_to_choose]
f20_sp <- f20_sp[cols_to_choose]
f19_sp <- f19_sp[cols_to_choose]
f18_sp <- f18_sp[cols_to_choose]
f17_sp <- f17_sp[cols_to_choose]
f16_sp <- f16_sp[cols_to_choose]
f15_sp <- f15_sp[cols_to_choose]
```

Preprocessing of data
```{r}
# Replace NAs by 0
f22_sp[is.na(f22_sp)] <- 0
f21_sp[is.na(f21_sp)] <- 0
f20_sp[is.na(f20_sp)] <- 0
f19_sp[is.na(f19_sp)] <- 0
f18_sp[is.na(f18_sp)] <- 0
f17_sp[is.na(f17_sp)] <- 0
f16_sp[is.na(f16_sp)] <- 0
f15_sp[is.na(f15_sp)] <- 0
# Group dataframes by team.
f22_gp <- f22_sp %>% group_by(club_name) %>% summarise(across(where(is.numeric), mean))
f21_gp <- f21_sp %>% group_by(club_name) %>% summarise(across(where(is.numeric), mean))
f20_gp <- f20_sp %>% group_by(club_name) %>% summarise(across(where(is.numeric), mean))
f19_gp <- f19_sp %>% group_by(club_name) %>% summarise(across(where(is.numeric), mean))
f18_gp <- f18_sp %>% group_by(club_name) %>% summarise(across(where(is.numeric), mean))
f17_gp <- f17_sp %>% group_by(club_name) %>% summarise(across(where(is.numeric), mean))
f16_gp <- f16_sp %>% group_by(club_name) %>% summarise(across(where(is.numeric), mean))
f15_gp <- f15_sp %>% group_by(club_name) %>% summarise(across(where(is.numeric), mean))
# Join dataframes
f22_gp <- left_join(f22_gp, table22, by = c("club_name"="Equipo"))
f21_gp <- left_join(f21_gp, table21, by = c("club_name"="Equipo"))
f20_gp <- left_join(f20_gp, table20, by = c("club_name"="Equipo"))
f19_gp <- left_join(f19_gp, table19, by = c("club_name"="Equipo"))
f18_gp <- left_join(f18_gp, table18, by = c("club_name"="Equipo"))
f17_gp <- left_join(f17_gp, table17, by = c("club_name"="Equipo"))
f16_gp <- left_join(f16_gp, table16, by = c("club_name"="Equipo"))
f15_gp <- left_join(f15_gp, table15, by = c("club_name"="Equipo"))
# Check for NAs in dataframe.
for(i in 1:length(cols_to_choose)){
  print(sum(is.na(f22_gp[i])))
}
```

Add a multiplier on points for second division teams going to first division.
```{r}
# second_div_mult <- 0.5
```

Create new dataframe
```{r}
new_df <- bind_rows(f21_gp, f20_gp, f19_gp, f18_gp, f17_gp, f16_gp, f15_gp)
```

Remove some columns
```{r}
colnames(new_df)
cols_to_drop <- c( "Pos.", "PJ", "G", "E", "P", "GF", "GC", "Dif.", "Des", "Asc")
new_df <- new_df[, !colnames(new_df) %in% cols_to_drop]
f22_gp <- f22_gp[, !colnames(f22_gp) %in% cols_to_drop]
```

TRAIN
```{r}
y <- new_df$Pts.
x <- select(new_df, -c("Pts.", "club_name") ) %>% as.matrix() %>% scale()
```

TEST 
#### NEEDS TO BE SAME SCALE!!
```{r}
y_te <- f22_gp[c("club_name","Pts.")]
x_te <- select(f22_gp, -c("Pts.", "club_name") ) %>% as.matrix() %>% scale()
```
# -----------------------------------
# ----------- LASSO MODEL -----------
# -----------------------------------
```{r}
fit.lasso= cv.glmnet(x=x, y=y, family = "gaussian", nfolds = 5, alpha=1, type.measure = "mse")

# #auc
lasso.err <- assess.glmnet(fit.lasso, newx=x, newy=y, family='gaussian')
lasso.err 

plot(fit.lasso$glmnet.fit, xvar='lambda')
plot(fit.lasso)

# try model w min value of lambda
b.lasso= as.vector(coef(fit.lasso, s='lambda.min'))
betas = data.frame("word" = c("intercept", colnames(x)), "beta_hat"= round(b.lasso, 5))
betas %>% arrange(desc(abs(beta_hat)))

# not as good but just to try something else
b.lasso= as.vector(coef(fit.lasso, s=0.1))
betas = data.frame("word" = c("intercept", colnames(x)), "beta_hat"= round(b.lasso, 5))
betas %>% arrange(desc(abs(beta_hat)))

# Test base model on 22 dataset
predictions <- predict(fit.lasso, newx = x_te)
y_te$pred <-predictions
y_te <- y_te %>% arrange(desc(Pts.))
```
```{r}
# Test other model on 22 dataset
predictions <- predict(fit.lasso, newx = x_te)
y_te$pred <-predictions
y_te <- y_te %>% arrange(desc(Pts.))
# For dataset
# compute averages by player
# calculate month of birth for player
# modify goalkeeper stats to 0 for 

# For model
# different values of lambda
```
# -----------------------------------
# --- BAYESIAN HIERARCHICAL MODEL ---
# -----------------------------------
```{r}
fp22_hm <- left_join(f22_sp, table22, by = c("club_name"="Equipo"))
fp21_hm <- left_join(f21_sp, table21, by = c("club_name"="Equipo"))
fp20_hm <- left_join(f20_sp, table20, by = c("club_name"="Equipo"))
fp19_hm <- left_join(f19_sp, table19, by = c("club_name"="Equipo"))
fp18_hm <- left_join(f18_sp, table18, by = c("club_name"="Equipo"))
fp17_hm <- left_join(f17_sp, table17, by = c("club_name"="Equipo"))
fp16_hm <- left_join(f16_sp, table16, by = c("club_name"="Equipo"))
fp15_hm <- left_join(f15_sp, table15, by = c("club_name"="Equipo"))


cols_to_drop <- c( "Pos.", "PJ", "G", "E", "P", "GF", "GC", "Dif.", "Des", "Asc")

y_te <- fp22_hm$Pts.
hm_data <-bind_rows(fp21_hm, fp20_hm, fp19_hm, fp18_hm, fp17_hm, fp16_hm, fp15_hm)
hm_data<- hm_data[, !colnames(hm_data) %in% cols_to_drop] 
fp22_hm<- fp22_hm[, !colnames(fp22_hm) %in% cols_to_drop] 

# fp22_hm[] <- lapply(fp22_hm, function(x) if(is.numeric(x)){
#   scale(x, center=TRUE, scale=TRUE)
#   } else x)

# Check for NAs in dataframe.
for(i in 1:length(cols_to_choose)){
  print(sum(is.na(f22_hm[i])))
}
fp22_hm$ran <- sample(-10:10, 633, rep = TRUE)
fp22_hm$ranpts <- fp22_hm$Pts.+ fp22_hm$ran


ML_ex <- lmer(Pts. ~ overall + value_eur + age + (1|club_name), data=hm_data, REML = FALSE) 
display(ML_ex)

c<-predict(ML_ex, newx=fp22_hm)
dim(hm_data)
rbind(y_te, predict(ML_ex, newx=fp22_hm))
length(y_te)
dim(predict(ML_ex, newx=fp22_hm))
dim(fp22_hm)

hm_data$pred <- predict(ML_ex, newx=hm_data)
hm_data[c("pred", "Pts.")]
cor(hm_data$pred, hm_data$Pts.)


grouped <- hm_data %>% group_by(club_name) %>% summarise(across(where(is.numeric), mean))
cor(grouped$pred, grouped$Pts.)

```

# -----------------------------------
# ---- BAYESIAN REGRESSION MODEL ----
# -----------------------------------

```{r}
# Using scaled_new_df dataframe for Bayesian regression model
scaled_new_df <- bind_cols(x, y)
names(scaled_new_df)[names(scaled_new_df) == '...54'] <- 'Pts.'
scaled_new_df <- select(scaled_new_df, -c(league_level))
```

```{r}
library(rstanarm)

# Fit the model:
model <- rstanarm::stan_glm(Pts. ~ ., family = gaussian(), data=scaled_new_df)

# View a summary of the model:
summary(model)
```

```{r}
# Making predictions:
x_te_df <- as.data.frame(x_te)
predictions_bayes <- predict(model, newdata = x_te_df)

pred_bayes_compare <- cbind(y_te, predictions_bayes)
pred_bayes_compare
```
Mean squared error to assess accuracy of Bayesian Regression
```{r}
rmse_bayes = mean((pred_bayes_compare$Pts. - pred_bayes_compare$predictions_bayes)^2)
rmse_bayes
```


# -----------------------------------
# ---- BAYESIAN MODEL SELECTION -----
# -----------------------------------
```{r}
library(mombf)
# Fit Bayesian model selection model
fit.bayesreg <- modelSelection(Pts. ~ ., data=scaled_new_df, priorCoef=zellnerprior(taustd=1), priorDelta=modelbbprior(1,1))

# View summary
head(postProb(fit.bayesreg),10)
```

# -----------------------------------
# --------- VISUALIZATIONS ----------
# -----------------------------------

```{r}
colnames(scaled_new_df)
```

```{r}
scaled_new_df[1, ]
```
```{r pressure, echo=FALSE}
library(ggplot2)
ggplot(data=new_df, aes(Pts.)) + geom_histogram(bins = 20) + ggtitle("Distribution of Points Scored")
#hist(new_df$Pts.)
```
```{r pressure, echo=FALSE}
ggplot(data=new_df, aes(Pts.)) + geom_density()
```

